---
title: "ewing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ewing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.heigth = 9,
  fig.width = 8.5
)
```

```{r setup}
library(ewing)
```

## Modular approach 

This was developed 18 Nov 2002 and updated 10 Aug 2022.

A big change is that I have modularized things. The default species are now 
"host" and "parasite". There is a second library called "[redscale](https://github.com/byandell/redscale)" that 
has redscale and aphytis and comperiella and encarsia. (Actually, host is 
just redscale, and parasite is just aphytis). Now you can set up a new 
system by just creating a new library package.

The second big change is that I use R conventions of passing arguments, 
rather than hiding things as I have done in the past. Now you do

```{r}
mysim <- init.simulation(interact = FALSE)
```

```{r}
simres <- future.events(mysim, refresh = 1000, plotit = TRUE)
```

```{r}
simres <- future.events(mysim, plotit = FALSE)
```

```{r}
ggplot_ewing(simres)
```

```{r}
ggplot_ewing(simres, total = FALSE, normalize = FALSE)
```

```{r}
ggplot_ewing(simres, normalize = FALSE, substrate = FALSE)
```

```{r}
plot(simres)
```

```{r}
plot(simres, substrate = FALSE)
```

The object "mysim" has the initial populations and the organism and 
temperature structures. The object "simres" has the populations after 
running future.events (used to be step.future), along with summary 
information from the run.

I have also modularized the immediate, pending and future events. The functions

	event.birth	birth immediate event
	event.attack	attack immediate event
	event.death	death pending event
	event.future	generic future event

You can supply your own version of these if you don't like mine. And you 
can add new ones by making appropriate changes in the "event" column of the 
future.host and future.parasite column. Adding a new kind of event, such as 
migrate means the routine now looks for a routine called "event.migrate".
Also, event.attack uses the type of parasite (endo,ecto) and the actual 
event (feed,ovip) to look for routines like event.ecto.feed to process 
attack details.

## Multiple Simulation Runs

```{r}
data(simdata)
```

```{r}
summary(simdata)
```

```{r}
ggplot_ewing_envelopes(simdata)
```

```{r}
ggplot_ewing_envelopes(simdata, confidence = TRUE)
```

## Other plots

```{r}
plot_current(simres, "host")
text_current(simres, "host")
```

```{r}
plot_current(simres, "parasite")
text_current(simres, "parasite")
```

The following plot is supposed to break out by substrate, which is in `sub.stage` in the `community` object (`simres`). But the way it is saved right now makes this goofy somehow.
Changing should be done carefully as this is used in multiple places. For now just leave it.
Note also that the color and text should correspond but not into the legend as I need
to formally assign the colors in a more intuitive way a la ggplot2. Later.

Need an explanation of single letter codes (say from future.host)

```{r}
ggplot_current(simres, "host")
```

Interactive plots showing splines. Use `graphics::locator()`.

```{r eval=FALSE}
five.plot()
```

```{r eval=FALSE}
five.show()
```

Plot to relate day to degree-day based on temperature regime.

```{r}
temp.plot(simres)
```

interactive design plot for high and low temperatures. Uses `graphics::locator()`.

```{r eval=FALSE}
temp.design(simres)
```

# Inner workings of ewing simulation

Simulations are driven by flat files that declare the species, their relationships with
each other and with the model substrate, and the details of scheduling future events for
individuals.
There are also files that specify the temperature regime, which is important if species
respond to degree-day (`DD`) cues rather than temporal cues; this aspect is only briefly
developed (see some plot ideas above) and will not be discussed further below.
Here we use the default files in the `data` part of this package.

These simulations keep track of the current stage and next scheduled future event for every individual. That is, these simulations are individual-based rather than population-based.
It is useful (such as for plots and other summaries) to collapse over individuals to get
population trends, but impossible to reverse this without unrealistic assumptions.
To that end, summary information at the population level (say by species life stages)
are accumulated during the simulation, and later simplified for plots and tables.

This simulation is set up so that names of species, their stages, and subtrates are
taken from the flat tables rather than being hardcoded. They are not limited to
two species, although ultimately interactions are dyadic (two individuals) and monadic
choices of future trajectories come down to competing risks for individuals.

## Initialization

One begins a simulation by initializing all species.
The `organism.features` table specifies the `units` (`hr` for hour or `DD` for degree-day),
`offspring` (either a Poisson mean number or the name of the `host`), species to `attack` (or `NA`), stage for `birth` (`gravid` for `host`), whether species relies on a `substrate`,
denominator to `deplete` individuals relative to time units that have passed,
which ageclass (`subclass`) of a species moves across the `substrate` (`host` for `host` and `adult`
for `parasite`; used for counting), what type is the parasite (`ecto` for `parasite` but could be `endo`),
and which stage of a species can actually `move` along the `substrate`.

Each species has a future event scheduling table (`future.host` and `future.parasite`).
The `init` column specifies the relative weight of `current` stages for simulating
the initial population.

## Future events and competing risks

Each row of a future event scheduling table has possible `future` events from a `current`
stage in the life of an individual of that species.
The `fid` points to the row in this table corresponding to the `future` stage.
The code uses numeric `fid` because it ends up in a vector of other numeric values.
An individual in a species will progress from `current` to `future` stage
when its event time is at the top of the event queue, unless some interaction leads to
a change for that individual. 
The `future` event time is scheduled using the `time` entry. When an individual
is at the top of the event queue, the `future` event time is scheduled using a spline
interpolation with mean from the `time` value. If the species uses hours (`hr`),
the spline is just a straight line, but if it uses `DD` the spline uses information
from the `temperature` profile for the simulation (again, not discussed here yet).

Note that sometimes there are multiple rows with the same `current` value, which are 
competing risks. For instance `future.host` has competing risks from the `current` stage
`second.3` of becoming `female` or `male`, while `future.parasite` has competing risk from
the current stage `adult` to `feed` or `ovip`osit, with return lines from `feed` and `ovip`
to `adult`. That is, an adult parasite might feed or oviposit, which have different health
and population consequences: feeding prolongs life while ovipositing produces new offspring
and depletes life.

All individuals have a `current` stage and are organized into an event queue,
which is a triply-linked leftist tree,
based on the scheduled time for their next `future` event.
Each species has its own leftist tree, with the tops of those trees
identifying the individuals with the closest (in time) next `future` event.
Internal routine `put.species`, in conjuction with
`leftist.update`, `leftist.remove` or `leftist.birth`, modify the leftist trees
when there is an individual event update, death (remove) or birth(s), respectfully.

## Plots and summaries

Plots used the plot character (`pch`), `color` (for stage) and `ageclass` (for summary
grouping). The `event` column is useful for summary tables and plotting, and sometimes
for other categorization (such as distinguishing `birth` and `death` from other types
of `future` events).

## Substrates and movement around triangular grid

This simulation system allows for individuals to be dispersed across a set of `substrate`s
and to move within and between `substrate` elements.
Currently, each `substrate` is designed on a triangular coordinate system of interconnected
triangles with diameter of 10 units (hardwired in internal routine `event.move`).
Connectivity of `substrate` segments is determined by table `substrate.substrate`, while
movement options for species are determined by species tables (`substrate.host` and
`substrate.parasite`). `substrate` elements may have sides (`1,2,3,4` for `fruit`, and
`top`, `bottom` for `leaf`). Relative weights for movement are specified for `init`ial
position, chance of a `parasite` to `find` a `host`, and choice to `move` among `substrate`
elements. The last columns of the `substrate` species tables give the relative weight
of moving from the current `substrate` to one of the other substrates.

The triangular grid and approximations used herein capture most (~95%) of movement
while simplifying calculations to max and min primarily rather than using quadratic
calculation needed for a rectangular coordinate system (where the Pythagorean theorem
rules).

## Redscale thoughts

The fields are mostly self-explanatory. DD represents mean degree days to that future event. Thus the mean time to "virgin" from "third.3" is 560-465=95 DD. The next two fields concern the Aphytis parasite, comprising the relative risk of being killed by an Aphytis that is feeding or is laying eggs, respectively. There was no explicit information on feeding risk, so these are made up following Luck's papers. The egg-laying risks are directly from the IPM pamphlet. The next two fields have the relative risks for the other two parasites. Then follows the gender (only females modeled here). The next field has the plotting symbol (pch), followed by the plotting color. The last and most important column is the life stage.

# Scaling up

The beginning simulations use hundreds of individuals and thousands of simulation steps,
as well as tens or hundreds of replicate simulations.
Theoretically, it should be possible to scale simulations up to hundreds of thousands
of individuals through some attention to loosely coupled systems and GPUs.
At small population sizes, these simulations may shed new light on shortcomings of
population-based simulations, but it is in large scale that they have the potential
to provide useful insights for systems-level study.

It is useful to note that the simulation steps through events rather than time.
Hence time is a function of events. Only in retrospect can the linear time
sequence, and population numbers, be measured. Further, the relationship between
steps and time is nonlinear, so that many steps may cover only a short time span,
or a few steps may traverse much time. Further, different species may have a
different relationship of steps to time. That is, the resolution for a species
is set by the granularity of events and their mean time duration. The span of
the simulation is determined by the number of allowed steps, and may result in
different spans for different species, when taking into account the unrealized
`future` events that remain in the community at the time of simulation stop.
