---
title: "envelope"
author: "Brian Yandell"
date: "2022-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 9, fig.height = 9)
```

```{r}
library(fda)
library(GET)
library(ggplot2)
library(patchwork)
```

# Girl heights with GET

See <https://cran.r-project.org/web/packages/GET/vignettes/GET.pdf>.

```{r}
years <- paste(1:18)
curves <- growth[['hgtf']][years,]
cset1 <- create_curve_set(list(r = as.numeric(years), 
                               obs = curves))
cset2 <- create_curve_set(list(r = as.numeric(years[-1]),
                               obs = curves[-1,] - curves[-nrow(curves),]))
```

Order  functions from the most extreme to the least extreme by the ’area’ measure

```{r}
A1 <- forder(cset1, measure = 'area'); order(A1)[1:10]
A2 <- forder(cset2, measure = 'area'); order(A2)[1:10]
```

Jointly order by combining these two by the ERL measure

```{r}
csets <- list(Height = cset1, Change = cset2)
A <- forder(csets, measure = 'area'); order(A)[1:10]
```

Highlight curves of the three most extreme girls. The plots of the two sets of
curves were produced using the GET and ggplot2 packages and combined by the patchwork
package (Pedersen 2020):

```{r}
cols <- c("#21908CFF", "#440154FF", "#5DC863FF")
p1 <- plot(cset1, idx = order(A)[1:3], col_idx = cols) + 
  labs(x = "Age (years)", y = "Height")
p2 <- plot(cset2, idx = order(A)[1:3], col_idx = cols) +
  labs(x = "Age (years)", y = "Change")
p1 + p2 + plot_layout(guides = "collect")
```

Use functional boxplot (Sun and Genton 2011) to see whether the most extreme girls are outliers with respect to height or its change.

```{r}
res <- fBoxplot(csets, type = 'area', factor = 1.5)
plot(res) + labs(x = "Age (years)", y = "Value")
```

# Interpolation of Ewing Curves


```{r}
library(ewing)
```

```{r}
if(file.exists(simfile <- "simfile.rds")) {
  out <- readRDS(simfile)
  nsim <- length(out)
} else {
  nsim <- 100
  out <- envelope_sim(nsim)
  saveRDS(out, file = simfile)
} 
```

```{r}
env <- ewing_envelope(out, "host", "time")
t1 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out, "parasite", "time")
t2 <- ggplot_ewing_envelope(env)
```

```{r}
t1 + t2 + plot_layout(nrow = 2)
```

Now try to do spline fit on each time step combination 

```{r}
out2 <- purrr::map(
  out,
  function(y) {
    purrr::map(
      y,
      function(x) {
        purrr::map_df(
          dplyr::distinct(
            purrr::map_df(
              dplyr::mutate(
                as.data.frame(x),
                time = ifelse(.data$step == 0, 0, ceiling(.data$time))),
              rev),
            .data$time, .keep_all = TRUE),
          rev)
      })})
```

```{r}
env <- ewing_envelope(out2, "host", "time", "step")
t1 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out2, "parasite", "time", "step")
t2 <- ggplot_ewing_envelope(env)
```

```{r}
t1 + t2 + plot_layout(nrow = 2)
```
## Host Counts over `r nsim` Simulations

Experimenting with time vs step

```{r}
env <- ewing_envelope(out2, "host", "crawler", "time")
p11 <- ggplot_ewing_envelope(env)
```

```{r}
p11
```

```{r}
env <- ewing_envelope(out, "host", "crawler")
p1 <- ggplot_ewing_envelope(env)
```

```{r}
p1
```

```{r}
env <- ewing_envelope(out, "host", "host")
p2 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out, "host", "gravid")
p3 <- ggplot_ewing_envelope(env)
```

```{r}
p1 + p2 + p3 + plot_layout(nrow = 3)
```

## Parasite Counts over `r nsim` Simulations

```{r}
env <- ewing_envelope(out, "parasite", "young")
p4 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out, "parasite", "adult")
p5 <- ggplot_ewing_envelope(env)
```

```{r}
p4 + p5 + plot_layout(nrow = 2)
```



