---
title: "envelope"
author: "Brian Yandell"
date: "2022-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, messages = FALSE, warnings = FALSE,
                      fig.width = 9, fig.height = 9)
```

Jim 20220925:
Assuming the “env” control the number of “simulations” how do I set up a scenario. Say I wanted to runs a series of simulations with the  number of hosts set at - say 100 , the number of parasites set at 1000 and  for 5000 iterations and calculate the envelope such a scenario? Does the Assuming I would like to run the “env’ routine, how would I do it?

Is this going to be available on the Population Ethology web site? If I wanted to look past means and standard deviations will envelope do that also? I have a lot to learn. I have started to look at the Isle Royale Biome study, but its going to take a while to figure out how convert that data into something that is useable for “ewing”.

I am looking forward to the “Start Simulation” button.

```{r}
library(patchwork)
```

# Interpolation of Ewing Curves


```{r}
library(ewing)
```

Conduct 100 simulations with default settings and save with discrete time increment of 0.5 units.

```{r}
if(file.exists(simfile <- "simfile.rds")) {
  out <- readRDS(simfile)
  nsim <- length(out)
} else {
  nsim <- 100
  out <- ewing_discrete(nsim, increment = 0.5)
  saveRDS(out, file = simfile)
} 
```

Simulation is done by `step`s, with `time` as a random, dependent variable. 
First we look at `time` as a function of `step`.

```{r}
env <- ewing_envelope(out, "host", "time", "step")
t1 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out, "parasite", "time", "step")
t2 <- ggplot_ewing_envelope(env)
```

```{r}
wrap_plots(t1, t2) + plot_layout(nrow = 2)
```

However, we saved simulations on discrete `time` steps of 0.5 units,
so we can look at `step` as a function of `time`.
Below all plots are with respect to time so that they are comparable across species.

```{r}
env <- ewing_envelope(out, "host", "step")
t1 <- ggplot_ewing_envelope(env)
```

```{r}
env <- ewing_envelope(out, "parasite", "step")
t2 <- ggplot_ewing_envelope(env)
```

```{r}
patchwork::wrap_plots(t1, t2) + patchwork::plot_layout(nrow = 2)
```

# Host and Parasite Counts over `r nsim` Simulations

```{r}
envs <- ewing_envelopes(out)
```

```{r}
summary(envs)
```

```{r}
ggplot_ewing_envelopes(out)
```

# Confidence Envelopes

```{r}
envs <- ewing_envelopes(out)
```

```{r}
ggplot_ewing_envelopes(envs, confidence = TRUE)
```

