---
title: "Ewing Data"
author: "Brian Yandell"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ewing)
library(tidyverse)
```

```{r}
data("organism.features")
data("future.host")
data("future.parasite")
data("host.parasite")
data("substrate.host")
data("substrate.parasite")
```

Get names of data structures.

```{r}
dlist <- c("future.host",
           "future.parasite",
           "host.parasite",
           "substrate.host",
           "substrate.parasite")
dlist <- as.list(dlist)
names(dlist) <- c(dlist)
nlist <- map(dlist, function(x) {
  names(get(x))
})
```

We can combine (rbind) the futures by adding a column for species.

```{r}
future <- bind_rows(
  host = future.host,
  parasite = future.parasite,
  .id = "species")
```

We can combine the substrate in a similar way, but first have to change the rownames into a column.

```{r}
ftemp <- function(x) {
  out <- data.frame(subname = rownames(x), x)
  rownames(out) <- NULL
  out
}
substrate <- bind_rows(
  host = ftemp(substrate.host),
  parasite = ftemp(substrate.parasite),
  .id = "species")

```

Then we have 4 types of files:

- future
- substrate
- host.parasite
- organism.features

The organism.features would work better as a long table.

```{r}
ftemp <- function(x, cc) {
  out <- sapply(x[cc,], as.character)
  names(out) <- NULL
  out
}
features <- bind_rows(
  host = data.frame(feature = colnames(organism.features),
                    value = ftemp(organism.features, 1)),
  parasite = data.frame(feature = colnames(organism.features),
                        value = ftemp(organism.features, 2)),
  .id = "species") %>%
  filter(!is.na(value))
```

The host.parasite could be appended to host with some adjustment.

- future.host has "starved" current state
- future.host has competing risks for "second.3" to "female" and "male"
- host.parasite has "parasite" current state
- host.parasite has rownames which should be "current"

Could do the following:

- add columns to future for "ovip" "feed" "offspring" "male"
- include NA entries for these for parasite
- add column to indicate what specied is parasite

## Translation

Need routines to translate between formats, or adjust routines
to handle this automatically.

